---
title: "Report of Course Project"
output:
  html_document: default
---

This is the report of Course Project in Pratical Machine Learning.

##Backgroud

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These types of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. 

With the data collected from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. Our purpose is to predict their manner in which they did in their exericse. 

##Data
The data can be downloaded from 

training data: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

testing data: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


Before we load the data into R, we can have a preview in excel. Many variables in these two files are nulls or NA. We believe that these variables should be removed from the predicotors.

The blank and NA should be interpreted as NA values when we load the data into R.
The code can be shown like below
```{r}
Data<-read.csv("pml-training.csv",na.strings=c("NA",""));
Prediction<-read.csv("pml-testing.csv",na.strings=c("NA",""));
dim(Data); dim(Prediction)
```

##Data Cleaning
We can not analysis the data before we remove these NA variables. Also, there will be some unrelevant variables, which should also be removed in our analysis.

* Remove these NA and unrelevant variables.
```{r}
CleaningData<-subset(Data[,colSums(is.na(Data))==0],select=-c(1:7))
Pre<-subset(Prediction[,colSums(is.na(Data))==0],select=-c(1:7))
dim(CleaningData); dim(Pre)
```
* Check the variables that have extremely low variance 
```{r}
library(caret)
zeroVar<-nearZeroVar(CleaningData[sapply(CleaningData,is.numeric)],saveMetrics=TRUE) 
pml<-CleaningData[,zeroVar$nzv==0];
dim(pml)
```
Now, we successfully reduced the data dimension from 160 to 53.

##Split data to training and testing for cross validation.
```{r}
set.seed(1)
inTrain<-createDataPartition(y=pml$classe,p=0.7,list=FALSE);
training<-pml[inTrain,];testing<-pml[-inTrain,];
dim(training);dim(testing)
```
We got 13737 samples and 52 variables for training, 5885 samples and 52 variables for testing. (The left one variable is the object "classe")

##Analysis
Noticed that the classe variable is a factor variable. We mainly used the "rpart" and "random forest" to predict the manner. 

* Rpart
```{r}
set.seed(1)
modelRpart<-train(training$classe~.,method="rpart",data=training);
modelRpart
confusionMatrix(testing$classe,predict(modelRpart,newdata=testing));
```

This method has a very low accuracy in predicting manner.

* Random Forest

This method use trees as building blocks to build more complex models. We use randomForest function directly which is much faster than train function.
```{r}
library(randomForest)
set.seed(1)
modelRF<-randomForest(classe~.,data=training,importance=TRUE);
modelRF
confusionMatrix(testing$classe,predict(modelRF,newdata=testing));

```

The accuracy in testing data is very closed to 100%.

```{r,fig.align='center'}
varImpPlot(modelRF)
```

we can see which variables have higher impact on the prediction.

With this two methods, we can know that the accuracy of random forest method is higher.

##Conclusion

Now, with the results showed before, we know that the random forest model did a good job. So we can predict the manner in testing data with random forest method.
```{r}
Prediction<-predict(modelRF,newdata=Pre)
Prediction
```






